Efficient FPGA Implementation of\\ Digit Parallel Online Arithmetic Operators

### 2014-05-18 FPT2014/bare_conf.tex
Add draft abstract

%\usepackage{cite}

@@

@@

% can use linebreaks \\ within to get better formatting as desired
\title{Efficient FPGA Implementation of\\ Overclocking Friendly Online Arithmetic Operators}

@@

\begin{abstract}
%\boldmath
The abstract goes here.
\end{abstract}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

@@

\subsection{Proposed Mapping Method} 
To tackle this problem, we first modify the structure of the online adder to enable an efficient FPGA mapping. The structure of a 3-digit online adder is given as an example in Fig.~\ref{Fig:ModifiedOA}. In this equivalent structure, the first FA in each 4:2 compressor is split into 2 parts, which only generate carry and sum respectively. In this case they can be mapped individually on 2 LUTs. The second FA, which generates the outputs, is unchanged and can be implemented using the fast-carry logic.

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDadder_FPGA_New.pdf}
	\caption{Modified structure of online adder. Left: an example of 3-digit online adder. The shaded part refers to the 2 logic blocks (LBs) that can be mapped onto 1 slice. Right: one LB. the dotted box outlines the logic that can be mapped onto 1 LUT and the corresponding fast-carry logic.}
	\label{Fig:ModifiedOA}
\end{figure}

@@

In this mapping approach, the theoretical area figures of the online adder in terms of the number of LUTs and Slices with respect to the operand word-lengths ($N$) are given in (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), respectively.
\begin{eqnarray}\label{Eq:OAareaLUT}
	Area\_LUT = 2N
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OAareaSlice}
	Area\_Slice = 1+\left\lceil\frac{N-2}{2}\right\rceil
\end{eqnarray}

\subsection{Performance Analysis}
In our experiments we compare 3 types of adders using Virtex-6 FPGAs: ripple carry adder (RCA), online adder which is implemented using behavioural description based on 4:2 compressors (OA\_behv) and online adder which is built using our proposed approach (OA\_real). Notice that in order to utilize the carry resources and to avoid logic optimizations by the synthesis tool, the Xilinx primitive component "Carry4" is used to create OA\_real.

The rated frequencies of all designs for a variety of operand word-lengths are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing in ISE 14.1. It can be seen that our approach outperforms RCA when $N>8$, and that our approach is always faster than the online adder based on 4:2 compressors. Also we can see that the frequency of RCA decreases drastically with respect to the increment of the operand word-length, whereas the online adder maintains a relatively high frequency even for large designs. This is because the critical path of the online adder is irrelevant to the operand word-lengths.

@@

The area comparisons are demonstrated in Fig.\ref{Fig:AdderArea}. The theoretical area of the online adder (OA\_theory) based on (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}) are also included. It can be seen that OA\_behv is not area efficient for FPGAs, as the dedicated FPGA resources are not fully utilized. In comparison, our implementation achieves significant area savings: $25\%\sim 33\%$ in LUTs and $67\%\sim77\%$ in Slices. The area overhead of our approach against RCA is $1.80\times\sim1.98\times$ 
in LUTs and $1.00\times\sim1.88\times$ in Slices. However this is close to the theoretical minimum area overhead for online adders.

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[LUTs]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_LUTs2.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[Slices]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_Slices2.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different binary adder implementations.}
  \label{Fig:AdderArea}
  %\vspace{-2ex}
\end{figure}

@@

However, in order to implement the digit parallel online multiplier (OM), large area budget is required because of the large area overhead of a single iteration stage and the overall stage numbers required. For instance in an $N$-digit OM, totally $(2N+\delta)$ iterations are required to generate $2N$ digits outputs, and $(N+\delta)$ iterations for the most significant $N$ digits result. Each iteration stage consists of all operations in Eq.xxx. In comparison, in an $N$-digit array multiplier there are only $(N-1)$ rows, each of which is basically an $N$-digit ripple-carry adder.

@@

\subsection{FPGA Implementation}
The structure diagram of a 4-digit OM using the proposed algorithm is shown in Fig.~\ref{Fig:PMStructure}(a). Generally in an $N$-digit OM, there are only $N$ stages to generate results with $2N$ digits. Each stage can be efficiently implemented using FPGAs, and the structure diagram of stage $j$ is shown in Fig.~\ref{Fig:PMStructure}(b). In each stage, an online adder is used to derive $W_{[j]}$. The selection logic, as described in Section xxx and Eq.xxx, takes 3 digits input (6 bits) and generate 1 digit output (2 bits). Hence it can be implemented using two 6-LUTs for each output bit. Similarly the generation of $P_{[j]}$ can also be implemented using one 6-LUT, since only the integer parts of $W_{[j]}$ might change due to the selection of $z_j$. In addition, the structure of Stage 1 can be further optimized by removing the online adder, because $P_{[0]}=0$.

@@

@@

Normally the OM is connected with other arithmetic operators in real applications. If the outputs of an OM is utilized for subsequent operations, only the most significant half of the products are required to maintain the consistency of word-length between inputs and outputs. In the conventional multiplier with standard binary arithmetic, this is achieved by either truncating or rounding the least significant half of the products. However both the computation time and the structure remains unchanged, because the results are generated from LSDs.

@@

\subsection{Performance Analysis}
In the experiments we compare 3 types of binary multipliers with different word-lengths. One is created from the Xilinx Core Generator [xxx] with standard arithmetic, and it is configured with speed optimization. The other two multipliers are with online arithmetic, but the results are generated with full precision and half precision, respectively.

@@

The comparison of area in terms of the number of LUTs and Slices with respect to different operand word-lengths are illustrated in Fig.xxx and Fig.xxx, respectively.

Although our proposed new online multiplier demonstrates a better operating frequency in the practice, we can push this one step further by operating the circuits beyond their deterministic region and allowing timing violations to happen. In this case, we examine the overclocking behaviour of all 3 types of multipliers in terms of the mean relative error (MRE), which is given in (xxx). For instance, Fig.xxx demonstrates the MRE of the 8-digit multipliers.

\section{Results}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

@@

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

@@

\bibliographystyle{./IEEEtran}
\bibliography{./IEEEabrv,./Reference.bib}

>>>>

\usepackage{cite}

@@

\usepackage{courier}	%use courier font, activate using \texttt{}

@@

% can use linebreaks \\ within to get better formatting as desired
\title{Efficient FPGA Implementation of\\ Digit Parallel Online Arithmetic Operators}

@@

\begin{abstract}
%\boldmath
Online arithmetic was designed for most significant digits first computing. It has been widely utilized for ASIC implementations with significant performance improvements. In addition, recent research has shown that digit parallel online operators can fail more gracefully when operating beyond the deterministic region, in comparison to the operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators requires large area overhead, and the efficient implementation of general purpose online operators on FPGAs remains under investigation. In this paper, we propose novel approaches to implement the online adder and the online multiplier efficiently on current FPGAs with 6-input LUTs. The FPGA carry resources are intensively used for both area reduction and performance improvements. The performance of the proposed architectures are experimentally demonstrated

%, they are targeting on FPGAs with 4-input LUTs, which are now gradually replaced by 6-input LUTs.

\end{abstract}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%\IEEEpeerreviewmaketitle

@@

\subsection{Proposed Mapping Method}
To tackle this problem, we first modify the structure of the online adder to enable an efficient FPGA mapping. The structure of a 4-digit online adder is given as an example in Fig.~\ref{Fig:ModifiedOA}. In this equivalent structure, the first FA in each 4:2 compressor is split into 2 parts, which only generate carry and sum respectively. In this case they can be mapped individually on 2 LUTs. The second FA, which generates the outputs, is unchanged and can be implemented using the fast-carry logic.

@@

\begin{figure*}[tbp]
	\centering
	\includegraphics[width=.85\textwidth]{./Figures/SDadder_FPGA_New_4digit.pdf}
	\caption{Modified structure of online adder. Left: an example of 4-digit online adder. The shaded part refers to the 2 logic blocks (LBs) that can be mapped onto 1 slice. Right: one LB. the dotted box outlines the logic that can be mapped onto 1 LUT and the corresponding fast-carry logic.}
	\label{Fig:ModifiedOA}
\end{figure*}

@@

In this mapping approach, the theoretical area figures of the online adder in terms of the number of LUTs and Slices with respect to the operand word-lengths ($N$) are given in (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), respectively.
\begin{eqnarray}\label{Eq:OAareaLUT}
	OAarea\_LUT = 2N
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OAareaSlice}
	OAarea\_Slice = 1+\left\lceil\frac{N-2}{2}\right\rceil
\end{eqnarray}

\subsection{Performance Analysis}
In our experiments we compare 3 types of adders using Virtex-6 FPGAs: ripple carry adder (\texttt{RCA}), online adder which is implemented using behavioural description based on 4:2 compressors (\texttt{OA\_behv}) and online adder which is built using our proposed approach (\texttt{OA\_new}). Notice that in order to utilize the carry resources and to avoid logic optimizations by the synthesis tool, the Xilinx primitive component "\texttt{Carry4}" is used to create \texttt{OA\_new}.

The rated frequencies of all designs for a variety of operand word-lengths are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing in ISE 14.1. It can be seen that our approach outperforms \texttt{RCA} when $N>8$, and that our approach is always faster than the online adder based on 4:2 compressors. Also we can see that the frequency of \texttt{RCA} decreases drastically with respect to the increment of the operand word-length, whereas the online adder maintains a relatively high frequency even for large designs. This is because the critical path of the online adder is irrelevant to the operand word-lengths.

@@

The area comparisons are demonstrated in Fig.\ref{Fig:AdderArea}. The theoretical area of the online adder (\texttt{OA\_theory}) based on (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}) are also included. It can be seen that \texttt{OA\_behv} is not area efficient for FPGAs, as the dedicated FPGA resources are not fully utilized. In comparison, our implementation achieves significant area savings: $25\%\sim 33\%$ in LUTs and $67\%\sim77\%$ in Slices. The area overhead of our approach against \texttt{RCA} is $1.80\times\sim1.98\times$
in LUTs and $1.00\times\sim1.88\times$ in Slices. However this is close to the theoretical minimum area overhead for online adders.

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[Usage of LUTs.]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_LUTs2.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[Usage of Slices]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_Slices2.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different binary adder implementations.}
  \label{Fig:AdderArea}
  %\vspace{-2ex}
\end{figure}

@@

However, in order to implement the digit parallel online multiplier (\texttt{OM}), large area budget is required because of the large area overhead of a single iteration stage and the overall stage numbers required. For instance in an $N$-digit \texttt{OM}, totally $(2N+\delta)$ iterations are required to generate $2N$ digits outputs, and $(N+\delta)$ iterations for the most significant $N$ digits result. Each iteration stage consists of all operations in Eq.xxx. In comparison, in an $N$-digit array multiplier there are only $(N-1)$ rows, each of which is basically an $N$-digit ripple-carry adder.

@@

\subsection{FPGA Implementation}
The structure diagram of a 4-digit \texttt{OM} using the proposed algorithm is shown in Fig.~\ref{Fig:PMStructure}(a). Generally in an $N$-digit \texttt{OM}, there are only $N$ stages to generate results with $2N$ digits. Each stage can be efficiently implemented using FPGAs, and the structure diagram of stage $j$ is shown in Fig.~\ref{Fig:PMStructure}(b). In each stage, an online adder is used to derive $W_{[j]}$. The selection logic, as described in Section xxx and Eq.xxx, takes 3 digits input (6 bits) and generate 1 digit output (2 bits). Hence it can be implemented using two 6-LUTs for each output bit. Similarly the generation of $P_{[j]}$ can also be implemented using one 6-LUT, since only the integer parts of $W_{[j]}$ might change due to the selection of $z_j$. In addition, the structure of Stage 1 can be further optimized by removing the online adder, because $P_{[0]}=0$.

@@

Notice that the blocks that generate partial products $Xy_j$ can be incorporated into the online adder for further area reduction. In the online adder as seen in Fig.~\ref{Fig:SliceNew}, maximally only 4 inputs per LUT are used. This leaves 2 available inputs per LUT. However in order to generate 1 digit partial product, 1 digit of inputs $X$ and $Y$ are required respectively, as indicated in Fig.~\ref{Fig:PartialProduct}. Therefore instead of using extra logic to generate $Xy_j$, it can be incorported into the corresponding LUTs in the online adder by fully utilizing all 6 LUT inputs.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.4\textwidth]{./Figures/PartialProduct.pdf}
	\caption{An example of achieving area reduction by combining the logic blocks that generate partial products into the Online Adder by fully utilizing the 6-LUTs.}
	\label{Fig:PartialProduct}
\end{figure}

The theoretical minimum resource usages of the proposed \texttt{OM} can be calculated as follows. In an $N$-digit \texttt{OM}, overall $N-1$ online adders are needed. According to Algorithm xxx, the word-length of each online adder is $N+2$-digit (2 integer digits and $N$ fractional digits). Therefore based on (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), the number of LUTs and the number of Slices used by the \texttt{OM} is given in (\ref{Eq:OMareaLUT}) and (\ref{Eq:OMareaSlice}), respectively. Note that $L_{S1}$ and $S_{S1}$ denote the LUT and Slice usages of Stage 1, which is purely built with combinational logic without online adders.

\begin{eqnarray}\label{Eq:OMareaLUT}
	OMarea\_{LUT} = 2(N+2)(N-1) + {L}_{S1}
\end{eqnarray}

\begin{eqnarray}\label{Eq:OMareaSlice}
	OMarea\_{Slice} = (1+\left\lceil{\frac{N}{2}}\right\rceil)(N-1)+S_{S1}
\end{eqnarray}

@@

Normally the \texttt{OM} is connected with other arithmetic operators in real applications. If the outputs of an \texttt{OM} is utilized for subsequent operations, only the most significant half of the products are required to maintain the consistency of word-length between inputs and outputs. In the conventional multiplier with standard binary arithmetic, this is achieved by either truncating or rounding the least significant half of the products. However both the computation time and the structure remains unchanged, because the results are generated from LSDs.

@@

\subsection{Performance Analysis}
In the experiments we compare 3 types of binary multipliers. One is created from the Xilinx Core Generator [xxx] with standard arithmetic. The \texttt{CoreGen} multiplier is implemented based on LUTs, and it is configured with speed optimization. The other two multipliers are both built with online arithmetic, except that the products are generated with full precision (\texttt{OM\_full}) and half precision (\texttt{OM\_half}), respectively.

@@

The comparison of area in terms of the number of LUTs and Slices with respect to different operand word-lengths are illustrated in Fig.~\ref{Fig:OMArea}. It can be seen that the area of \texttt{OM\_full} is close to the theoretical values predicted by (\ref{Eq:OMareaLUT}) and (\ref{Eq:OMareaSlice}). In comparison to the \texttt{CoreGen} multiplier, the area overheads of \texttt{OM\_full} are $48\%\sim84\%$ for LUTs, and $0\%\sim83\%$ for Slices. On the other hand, the \texttt{OM\_half} achieves area reductions for all operand word-lengths when compared to the \texttt{CoreGen} multiplier. The area saving varies from $40\%$ to $55\%$ for LUTs, and $46\%\sim55\%$ for Slices.

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[Usage of LUTs.]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/OM_LUTs.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[Usage of Slices]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/OM_Slice.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different binary multipliers.}
  \label{Fig:OMArea}
  %\vspace{-2ex}
\end{figure}

Although our proposed new online multiplier demonstrates a better operating frequency in the practice, we can push this one step further by operating the circuits beyond their deterministic region and allowing timing violations to happen. In this case, we examine the overclocking behavior of all 3 types of multipliers in terms of the mean relative error (MRE), which is given in (xxx). For instance, Fig.xxx demonstrates the MRE of the 8-digit multipliers.\cite{Ercegovac_DSPOnline}

%\section{Results}

%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

@@

@@

\bibliographystyle{./IEEEtran}

% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}

\bibliography{./IEEEabrv,./Reference}
